---
title: "PISA"
output: html_notebook
---

Loading data:
```{r}
#packages
library(readxl)
library(MASS)
library(car)
library(ggplot2)
library(data.table)
library(psych)
library(simpleboot)
library(psychometric)
library(boot)
library(corrplot)
library(RColorBrewer)
library(reshape2)

#loading data
PISA <- read_excel("PISA.xlsx")
View(PISA)

#renaming columns for later visualisations
colnames(PISA)[2] <- "Age"
colnames(PISA)[6] <- "Emotion"

#Creating total ToM scores (Cognitive and Affective Theory of Mind)
PISA$Cog_ToM <- PISA$do + PISA$think
PISA$Aff_ToM <- PISA$feel
```

Making summary statistics:
```{r}
#total summary
summary(PISA)

#total proportion of males to females
prop.table(table(PISA$sex))
```

Correlation Matrix

```{r}
PISACor <- data.frame(PISA$happy, PISA$fear, PISA$disgust, PISA$anger, PISA$surprise, PISA$sad, PISA$neutral, PISA$Emotion, PISA$do, PISA$think, PISA$feel, PISA$sincere, PISA$sarcastic, PISA$Cog_ToM, PISA$EC, PISA$PD, PISA$EE)

#spearman values
res1 <- corr.test(PISACor, method = "spearman", use = "complete.obs")$r
res1

#p-values
res2 <- corr.test(PISACor, method = "spearman", use = "complete.obs")$p
res2
```


Running multiple linear regressions with control variables:
```{r}
#Happy
multiHappy <- lm(happy ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiHappy)

#Happy
multiFear <- lm(fear ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiFear)

#Disgust
multiDisgust <- lm(disgust ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiDisgust)

#Anger
multiAnger <- lm(anger ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiAnger)

#Surprise
multiSurprise <- lm(surprise ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiSurprise)

#Sad
multiSad <- lm(sad ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiSad)

#Neutral
multiNeutral <- lm(neutral ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiNeutral)

#Emotion
multiEmotion <- lm(Emotion ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiEmotion)

#Feel
multiFeel <- lm(feel ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiFeel)

#Think
multiThink <- lm(think ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiThink)

#Do
multiDo <- lm(do ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiDo)

#Sincere
multiSincere <- lm(sincere ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiSincere)

#Sarcastic
multiSarcastic <- lm(sarcastic ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiSarcastic)

#Cog_ToM
multiCogToM <- lm(Cog_ToM ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiCogToM)

#EC
multiEC <- lm(EC ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiEC)

#PD
multiPD <- lm(PD ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiPD)

#EE
multiEE <- lm(EE ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA)
summary(multiEE)
```



Checking linearity assumption:
```{r}
#Happy
plot(multiHappy, 1)

#Fear
plot(multiFear, 1)

#Disgust
plot(multiDisgust, 1)

#Anger
plot(multiAnger, 1)

#Surprise
plot(multiSurprise, 1)

#Sad
plot(multiSad, 1)

#Neutral
plot(multiNeutral, 1)

#Emotion
plot(multiEmotion, 1)

#Feel
plot(multiFeel, 1)

#Think
plot(multiThink, 1)

#Do
plot(multiDo, 1)

#Sincere
plot(multiSincere, 1)

#Sarcastic
plot(multiSarcastic, 1)

#Cog_ToM
plot(multiCogToM, 1)

#EC
plot(multiEC, 1)

#PD
plot(multiPD, 1)

#EE
plot(multiEE, 1)

```



Checking normality assumption with QQ plots:
```{r}
#Happy
plot(multiHappy, 2)

#Fear
plot(multiFear, 2)

#Disgust
plot(multiDisgust, 2)

#Anger
plot(multiAnger, 2)

#Surprise
plot(multiSurprise, 2)

#Sad
plot(multiSad, 2)

#Neutral
plot(multiNeutral, 2)

#Emotion
plot(multiEmotion, 2)

#Feel
plot(multiFeel, 2)

#Think
plot(multiThink, 2)

#Do
plot(multiDo, 2)

#Sincere
plot(multiSincere, 2)

#Sarcastic
plot(multiSarcastic, 2)

#Cog_ToM
plot(multiCogToM, 2)

#EC
plot(multiEC, 2)

#PD
plot(multiPD, 2)

#EE
plot(multiEE, 2)

```


Checking normality with Shapiro-Wilk:
```{r}
#Happy
sresidMultiHappy <- studres(multiHappy)
shapiro.test(sresidMultiHappy)

#Fear
sresidMultiFear <- studres(multiFear)
shapiro.test(sresidMultiFear)

#Disgust
sresidMultiDisgust <- studres(multiDisgust)
shapiro.test(sresidMultiDisgust)

#Anger
sresidMultiAnger <- studres(multiAnger)
shapiro.test(sresidMultiAnger)

#Surprise
sresidMultiSurprise <- studres(multiSurprise)
shapiro.test(sresidMultiSurprise)

#Sad
sresidMultiSad <- studres(multiSad)
shapiro.test(sresidMultiSad)

#Neutral
sresidMultiNeutral <- studres(multiNeutral)
shapiro.test(sresidMultiNeutral)

#Emotion
sresidMultiEmotion <- studres(multiEmotion)
shapiro.test(sresidMultiEmotion)

#Feel
sresidMultiFeel <- studres(multiFeel)
shapiro.test(sresidMultiFeel)

#Think
sresidMultiThink <- studres(multiThink)
shapiro.test(sresidMultiThink)

#Do
sresidMultiDo <- studres(multiDo)
shapiro.test(sresidMultiDo)

#Sincere
sresidMultiSincere <- studres(multiSincere)
shapiro.test(sresidMultiSincere)

#Sarcastic
sresidMultiSarcastic <- studres(multiSarcastic)
shapiro.test(sresidMultiSarcastic)

#Cog_ToM
sresidMultiToMCoG <- studres(multiCogToM)
shapiro.test(sresidMultiToMCoG)

#EC
sresidMultiEC <- studres(multiEC)
shapiro.test(sresidMultiEC)

#PD
sresidMultiPD <- studres(multiPD)
shapiro.test(sresidMultiPD)

#EE
sresidMultiEE <- studres(multiEE)
shapiro.test(sresidMultiEE)

```
The residuals of all of models are non-normal except for Emotion, EC and PD. 

Checking the homoscedasticity assumption with scale-location plots:
```{r}

#Emotion
plot(multiEmotion, 3)

#EC
plot(multiEC, 3)

#PD
plot(multiPD, 3)

```

Checking homoscedasticity with ncvTest():
```{r}

#Emotion
ncvTest(multiEmotion)

#EC
ncvTest(multiEC)

#PD
ncvTest(multiPD)

```
Homoscedasticity assumption holds for all.

Bootstrapping to account for non-normality among residuals:
```{r}
set.seed(3333)
#Happy
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(happy ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)
b$t0 
b
boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(happy ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(happy ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)

#Fear
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(fear ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(fear ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(fear ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Disgust
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(disgust ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(disgust ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(disgust ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Anger
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(anger ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(anger ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(anger ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Surprise
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(surprise ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(surprise ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(surprise ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Sad
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(sad ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(sad ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(sad ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Neutral
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(neutral ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(neutral ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(neutral ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Emotion
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(Emotion ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(Emotion ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(Emotion ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Do
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(do ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(do ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(do ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Think
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(think ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(think ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(think ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Feel
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(feel ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(feel ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(feel ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Sincere
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(sincere ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(sincere ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(sincere ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#Sarcastic
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(sarcastic ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(sarcastic ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(sarcastic ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)

#CogTom
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(Cog_ToM ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(Cog_ToM ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(Cog_ToM ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#EC
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(EC ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(EC ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(EC ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#PD
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(PD ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(PD ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(PD ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)


#EE
#Finding 95%CI
boot.reg <- function(data, idx) {
  fit <- lm(EE ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coef(fit)
}

b <- boot::boot(PISA, boot.reg, 2000)


boot.ci(b, index = 2, type = "bca")
boot.ci(b, index = 3, type = "bca")
boot.ci(b, index = 4, type = "bca")
boot.ci(b, index = 5, type = "bca")
boot.ci(b, index = 6, type = "bca")
boot.ci(b, index = 7, type = "bca")

#Finding p-values
boot.reg <- function(data, idx) {
  fit <- lm(EE ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$coefficients[,4]
}

b <- boot::boot(PISA, boot.reg, 2000)

b

mean(b$t[,2])
mean(b$t[,3])
mean(b$t[,4])
mean(b$t[,5])
mean(b$t[,6])
mean(b$t[,7])

#Finding R2
boot.reg <- function(data, idx) {
  fit <- lm(EE ~ Age + sex + education + NARTIQ + HADSa + HADSd, data = PISA[idx,])
  coefs<-coef(fit)
  coefs
  summary(fit)$adj.r.squared
}

b <- boot::boot(PISA, boot.reg, 2000)

mean(b$t)

```

Standardising PISA and running bootstrapped regressions to collect Beta values:
```{r}
#standardising all scores except for predictor variables
PISAStandard <- PISA %>%
  dplyr::select(-ID, -Age, -sex, -education, -NARTIQ, -HADSa, -HADSd)

PISAnew <- PISA %>%
  dplyr::select(ID, Age, sex, education, NARTIQ, HADSa, HADSd)

PISAStandard <- lapply(PISAStandard, scale)

PISAStandard <- cbind(PISAStandard, PISAnew)

#bootstrapping to collect beta values.
set.seed(3333)

#Happy
happyStandard <- lm(happy ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(happyStandardBoot <- lm.boot(happyStandard, R = 2000))

#Fear
fearStandard <- lm(fear ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(fearStandardBoot <- lm.boot(fearStandard, R = 2000))

#Disgust
disgustStandard <- lm(disgust ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(disgustStandardBoot <- lm.boot(disgustStandard, R = 2000))

#Anger
angerStandard <- lm(anger ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(angerStandardBoot <- lm.boot(angerStandard, R = 2000))

#Surprise
surpriseStandard <- lm(surprise ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(surpriseStandardBoot <- lm.boot(surpriseStandard, R = 2000))

#Sad
sadStandard <- lm(sad ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(sadStandardBoot <- lm.boot(sadStandard, R = 2000))

#Neutral
neutralStandard <- lm(neutral ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(neutralStandardBoot <- lm.boot(neutralStandard, R = 2000))

#Emotion
EmotionStandard <- lm(Emotion ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(EmotionStandardBoot <- lm.boot(EmotionStandard, R = 2000))

#Do
doStandard <- lm(do ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(doStandardBoot <- lm.boot(doStandard, R = 2000))

#Think
thinkStandard <- lm(think ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(thinkStandardBoot <- lm.boot(thinkStandard, R = 2000))

#Feel
feelStandard <- lm(feel ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(feelStandardBoot <- lm.boot(feelStandard, R = 2000))

#Sincere
sincereStandard <- lm(sincere ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(sincereStandardBoot <- lm.boot(sincereStandard, R = 2000))

#Sarcastic
sarcasticStandard <- lm(sarcastic ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(sarcasticStandardBoot <- lm.boot(sarcasticStandard, R = 2000))

#Cognitive Theory of Mind
CogTomStandard <- lm(Cog_ToM ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(CogTomStandardBoot <- lm.boot(CogTomStandard, R = 2000))

#Empathic Concern
ECStandard <- lm(EC ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(ECStandardBoot <- lm.boot(ECStandard, R = 2000))

#Personal Distress
PDStandard <- lm(PD ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(PDStandardBoot <- lm.boot(PDStandard, R = 2000))

#Emotional Empathy
EEStandard <- lm(EE ~ Age + sex + education + NARTIQ + HADSa + HADSd, PISAStandard)
(EEStandardBoot <- lm.boot(EEStandard, R = 2000))
```

Heatmap of standardised coefficients:
```{r}
#Setting out all of the beta values
beta1 <- c(-0.007321,0.070210,-0.011867,0.005283,-0.002919,-0.007833)
beta2 <- c(-0.007321,0.070210,-0.011867,0.005283,-0.002919,-0.007833)
beta3 <- c(0.010632,0.521656,-0.002788,0.017886,-0.009282,-0.005646)
beta4 <- c(-0.022571,0.409137,-0.008287,0.030422,0.005430,-0.014098)
beta5 <- c(-0.021399,0.393212,-0.035219,0.015616,0.007336,-0.028634)
beta6 <- c(0.0061321,0.4596303,0.0026954,0.0228639,-0.0143485,-0.0003229)
beta7 <- c(-0.011866,-0.029460,-0.074377,0.032756,0.003097,-0.041971)
beta8 <- c(-0.01545,0.70356,0.01122,0.03785,-0.01507,-0.01325)
beta9 <- c(-0.04474,0.26021,-0.06434,0.01168,-0.01437,-0.02117)
beta10 <- c(-0.034978,0.060663,-0.038859,0.014658,0.004512,-0.030183)
beta11 <- c(-0.043546,-0.123750,0.009111,0.007482,-0.014008,0.002636)
beta12 <- c(-0.042034,0.025543,-0.015439,-0.002917,-0.016124,-0.026639)
beta13 <- c(-0.02559,0.13511,-0.05909,0.03687,0.01329,0.00421)
beta14 <- c(-0.043897, 0.174853, -0.056679, 0.014576, -0.005228, -0.028450)
beta15 <- c(0.012299,0.492174,-0.019322,0.008077,0.051126,-0.072041)
beta16 <- c(0.010968,0.202151,-0.016857,-0.006822,0.063182,0.071792)
beta17 <- c(1.807e-02,5.271e-01,-2.400e-02,-7.827e-05,8.733e-02,-3.971e-03)

#compiling into dataframe and heatmap
heatmap <- data.frame(beta1, beta2, beta3, beta4, beta5, beta6, beta7, beta8, beta9, beta10, beta11, beta12, beta13, beta14, beta15, beta16, beta17)
heatmap <- as.matrix(heatmap)
heatmap <- t(heatmap)
colnames(heatmap)<- c("Age", "Sex", "Education", "NARTIQ", "HADSa", "HADSd") 
rownames(heatmap)<- c("Happy", "Fear", "Disgust", "Anger", "Surprise", "Sad", "Neutral", "EmotionTotal", "Do", "Think", "Feel", "Sincere", "Sarcastic", "Cognitive ToM", "Empathic Concern", "Personal Distress", "Emotional Empathy")
heatmap <- round(heatmap,3)
heatmap

#setting out all of the p-values
p1 <- c(0.496, 0.468, .577, .455, .432, .458)
p2 <- c(.219, .276, .091, .327, .435, .470)
p3 <- c(.365, .037, .505, .179, .460, .473)
p4 <- c(.108, .072, .501, .022, .454, .440)
p5 <- c(.153, .082, .295, .211, .477, .394)
p6 <- c(.442, .048, .500, .074, .421, .476)
p7 <- c(.332, .477, .057, .017, .553, .292)
p8 <- c(.229, .001, .494, .003, .436, .488)
p9 <- c(.004, .224, .104, .323, .452, .424)
p10 <- c(.027, .495, .306, .270, .488, .370)
p11 <- c(.004, .433, .498, .439, .436, .414)
p12 <- c(.007, .516, .472, .529, .449, .381)
p13 <- c(.094, .412, .123, .018, .448, .533)
p14 <- c(.006, .356, .141, .268, .491, .384)
p15 <- c(.337, .047, .434, .414, .171, .121)
p16 <- c(.363, .336, .444, .446, .080, .120)
p17 <- c(.203, .030, .410, .514, .023, .517)

#compiling into dataframe
p_map <- data.frame(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13, p14, p15, p16, p17)
p_map <- as.matrix(p_map)
p_map <- t(p_map)
colnames(p_map)<- c("Age", "Sex", "Education", "NARTIQ", "HADSa", "HADSd")
rownames(p_map)<- c("Happy", "Fear", "Disgust", "Anger", "Surprise", "Sad", "Neutral", "EmotionTotal", "Do", "Think", "Feel", "Sincere", "Sarcastic", "Cognitive ToM", "Empathic Concern", "Personal Distress", "Emotional Empathy") 

#mergining the two dataframes to create final heatmap.
tiff(filename = "test8.tiff", compression = "lzw", width = 7, height = 10, units = "in", res = 300)
heatmapfig <- corrplot(heatmap, col=brewer.pal(n=8, name="RdBu"), tl.col = "black", tl.srt = 60, cl.ratio = 0.4, p.mat = p_map, sig.level = 0.017, insig = "pch", pch.col = "gray49", pch.cex = 2, method = "color", tl.cex = 1.2, cl.cex = 1.1, addgrid.col = "black")
dev.off()
```

Visualising betas:
```{r}
#creating dataframe comprising the four total scores against age
betaComp <- data.frame(Age = PISA$Age, Emotion = PISA$Emotion, Cog_ToM = PISA$Cog_ToM, Aff_ToM=PISA$Aff_ToM, EE = PISA$EE)
betaComp <- melt(setDT(betaComp), id.vars = c("Age"), variable.name = "domain")

#creating labels for the four beta values
lbls <- data.frame(domain = unique(betaComp$domain), labels = c("B = -.05", "B = -.10", "B = -.05","B = .11"), face = c("plain", "bold", "bold", "plain"))

#establishing predicted values for lines and CI
ages <- 40:82
ages <- rep(ages, 4)

#predicted values at each age level, for use in generating regression lines.
emotionvals <- ages[1:43] * -0.05 + 29.78040
cogvals <- ages[1:43] * -0.10 + 20.39160
affvals <- ages[1:43] * -0.05 + 10.42732
eevals <- ages[1:43] * 0.11 + 24.52901
values <- c(emotionvals, cogvals, affvals, eevals)

#minimum values, representing the lower bounds of the 95%CI, for use in error shading.
emotionmin <- ages[1:43] * -0.11 + 29.78040
cogmin <- ages[1:43] * -0.14 + 20.39160
affmin <- ages[1:43] * -0.07 + 10.42732
eemin <- ages[1:43] * -0.01 + 24.52901
minimum <- c(emotionmin, cogmin, affmin, eemin)

#maximum values, representing the upper bounds of the 95%CI, for use in error shading.
emotionmax <- ages[1:43] * 0.02 + 29.78040
cogmax <- ages[1:43] * -0.04 + 20.39160
affmax <- ages[1:43] * -0.02 + 10.42732
eemax <- ages[1:43] * 0.25 + 24.52901
maximum <- c(emotionmax, cogmax, affmax, eemax)

domain <- c(rep("Emotion", 43), rep("Cog_ToM", 43), rep("Aff_ToM", 43), rep("EE", 43))

line <- data.frame(ages, domain, values, minimum, maximum)

betaComp$domain <- factor(betaComp$domain, levels = c("Emotion", "Cog_ToM", "Aff_ToM", "EE"))
line$domain <- factor(line$domain, levels = c("Emotion", "Cog_ToM", "Aff_ToM", "EE"))

#generating final figure
tiff(filename = "test9.tiff", width = 5, height = 5, units = "in", res = 250)
ggplot(betaComp, aes(x = Age, y = value)) + 
  geom_jitter() + 
  geom_line(data = line, aes(x = ages, y = values), col = "Red", size = 1) + 
  geom_ribbon(data = line, aes(x = ages, ymin = minimum, ymax = maximum), inherit.aes = FALSE, alpha = 0.2) +
  #geom_abline(data = line, aes(intercept = intercepts, slope = min), linetype = "dashed", col = "grey") +
  #geom_abline(data = line, aes(intercept = intercepts, slope = max), linetype = "dashed", col = "grey") +
  facet_wrap(vars(domain), ncol = 2, labeller = labeller(domain = 
    c("Emotion" = "Emotion",
      "Cog_ToM" = "Cognitive ToM",
      "Aff_ToM" = "Affective ToM",
      "EE" = "Emotional Empathy"))) +
  xlab("Age") +
  ylab("Value") +
  geom_text(data = lbls, aes(label = labels, x = 71, y = 54, hjust = 0, vjust = 1, fontface = face))
dev.off()
```
